Iterative forecasting and the trap of self-fulfilling prophecies

### 2021-04-02 manuscript/manuscript.Rmd
:pen:

Let us assume $\xi_t(\sigma)$ represents log-normal random noise with mean of unity and log-standard-deviation $\sigma_i$ for each model.
Model 1 is given by $r_1 = 2$, $K_1 = 16$, $\sigma_1 = 0.05$, Model 2 by $r_2 = 0.5$,  $K_2 = 10$, $\sigma = 0.075$ (in dimensionless units).
Having both the larger growth rate and the larger carrying capacity, Model 1 is clearly the more optimistic of the two choices.

Selecting between Model 1 and Model 2 can thus be considered the simplest illustration of the model uncertainty problem.
This is a subset of the more general problem of selecting model parameters, assuming a logistic growth, which itself is a subset of estimating the best structural from (e.g. Ricker, Beverton-Holt, etc).
There is no need to consider these more complicated versions of the model uncertainty problem here, since they all inherit the same issue.
Reducing the model selection problem to these two models simplifies the presentation and will aid intuition at no loss of generality.

@@

The task of deciding whether Model 1 or Model 2 would be the better choice for decision making is thus perhaps the simplest example of the much studied issue of model uncertainty that we can pose. 
As in any real world scenario, neither model is the true model, but nevertheless this model set contains a good enough approximation of the true model to make good decisions.
However, any of the well-developed approaches for decision-making under model uncertainty will prefer Model 2 over Model 1, despite the fact that the optimal policy under Model 2 leads to much worse outcomes ecologically and economically.

@@

I illustrate the process of model selection by strictly proper scoring rules using two scenarios.
In Scenario A (passive observation) the fish stock is unharvested and allowed to recover towards carrying capacity (as simulated under our "true" model, Model 3) while comparing the observed stock size in each subsequent time step to the distribution predicted under Model 1 and Model 2 respectively [Fig 1].
The mean, $\mu_t$ and variance, $\sigma_t$ of the forecast are compared against the true observation $x_t$ using a proper scoring rule given by @Gneiting2007,

@@

In Scenario B (actively harvest), I have first solved for the optimal management strategy for both Model 1 and Model 2 [Fig 1b].
For small noise and concave functions with linear reward structure this can be done analytically [see proof in @Reed1979], or solved more generally by stochastic dynamic programming [see review by @Marescot2013, details in Appendix].
Under this scenario, replicate simulations of the stock are harvested at each time step using the optimal quota dictated by either Model 1 and Model 2.
The resulting stock sizes in the time-step following this harvest are once again compared to the probabilities predicted by each model using Eq \eqref{proper}.
Model 2 unequivocally outperforms Model 1 in both scenarios of passive observation and active harvest.

Despite the clearly superior predictive accuracy of Model 2 in both scenarios, the outcomes from management under Model 2 are substantially worse.
We can assess such outcomes in less abstract terms than forecasting skill, such as economic value (in dollars) or the ecological value (unharvested biomass).
In our simple formulation of the decision problem, the "utility" the manager seeks to maximize is simply the economic value (net present value: the discounted sum of all profits from future harvests, Eq \eqref{utility}) of harvested fish.
This formulation ignores any utility provided by fish that are not harvested, beyond their contribution to future potential harvests.
While it is possible to include such contributions directly in the utility function being optimized [e.g. @Halpern2013], even without doing so, Model 1 maintains both a higher unharvested biomass and also leads to higher economic returns throughout [Fig 2].

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from Model 1 result in a significantly higher fish stock size than under Model 2 (panel A). Economic returns under Model 1 are also substantially higher (panel B)"}

@@

In both scenarios, the careful comparison through proper scoring rules has led us to select the worse-performing model.
Crucially, a manager operating under this selection would have little indication that their model was flawed: both future stock sizes and expected harvest yields consistently match model predictions.
Had we been able to include Model 3 in our forecast comparisons, it would equal or outperform the forecasting skill of both Model 1 and Model 2 (as guaranteed by the theorem of @Gneiting2007), while also matching or out-performing their economic utility (as guaranteed by the theorem of @Reed1979).
In practice, we never have access to the generating model, so it is reasonable to expect model selection to determine the best approximation.
As we see here, the best approximation for forecasting future states does not in fact lead to better outcomes.

@@

Any adaptive management strategy updates posterior distributions over model uncertainty [@Punt2016; @Ludwig1982].
Unfortunately, any such adaptive updating leads to worse outcomes than the equivalent non-adaptive strategy, in which model uncertainty is held fixed.
I illustrate the application of a passive adaptive management strategy to this simple example, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty.
Passive adaptive management for a simple sequential decision problem is straight forward to implement over a discrete set of states and actions using dynamic programming with iterative updates [@Smith1981, example code in Appendix].
To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.
After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct  [Fig 3A].
I compare these results to 
As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating, which achieves a net present value that only 31% that expected under management using Model 1 alone [Fig 2].

```{r figure3, fig.height=5, fig.width=8, fig.cap="Adaptive management under model uncertainty.  Solid lines trace the trajectories of the state (fish stock, circles) and action (harvest quota, triangles), under adaptive management (learning).  Dotted lines trace the corresponding trajectories if iterative learning is omitted, leaving the prior belief fixed throughout the simulation (planning).   Color indicates the belief that Model 1 is correct (blue), with an intial prior belief of 99%.  Panel A: Management over the two candidate models, Model 1 and Model 2. Within a single iteration of adaptive management, the belief over models switches from a prior belief that heavily favored Model 1 to a posterior that favors Model 2 with near certainty.  Future iterations reinfornce the belief in Model 2, resulting in both depressed harvests and low stock sizes (solid lines).  If no iterative learning updates are performed, stock sizes and realized harvests (and thus economic profit) are both higher. Panel B: given 42 candidate models over a broad range of parameters, adaptive management quickly reduces the probability on model 1, and substantially underperforms management without learning (dotted lines), though not over-harvesting quite as badly as under the two-model scenario."}
am <- read_csv("../data/am.csv")
fig3a <- am %>%  filter(time <=20) %>% 
  pivot_longer(c(stock, quota), values_to="stock", names_to="key") %>%
  mutate(key = factor(key, levels = c("stock", "quota"))) %>%
  ggplot(aes(time, stock, shape = key, group = interaction(method,key))) + 
  geom_line(aes(lty=method), col = "gray40", show.legend = FALSE) + 
  geom_point(aes(col = belief), size=2, show.legend = FALSE) +
  scale_colour_gradient(limits = c(0,1), low = pal[4], high = pal[1]) +
  ylab("fish stock") + labs(subtitle="A.")

@@

Increasing the space of possible models to cover a whole plausible range of parameters $r$ and $K$ does not resolve this problem.
Learning is somewhat slower over the larger range of possibilities, but nevertheless converges towards parameter values with lower growth rates and lower carrying capacity, for which the target biomass is far lower [Fig 3B; details in appendix B]. The manager becomes trapped in vicious cycle as biomass declines further and further, subsequent observations seem more and more consistent with models which most favor over-fishing in the first place.  As biomass declines, annual harvests and profits also plunge, and yet our manager becomes only more confident in their decisions.

@@

<!-- quick summary of what was demonstrated -->
Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other. Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.  In both approaches, the manager quickly concludes that Model 1 is entirely implausible, while finding that Model 2 is remarkably accurate at predicting future values. This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models with all possible $K$ and $r$ values): in fact leads to a model with slightly lower $K$ and even higher level of over-fishing and under-performance than Model 2 (Fig 3B).

<!-- Intuition, Figure 4 -->
The reason for Model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, Model 3.  Plotting the growth rate functions of each model, [Fig 4A], it is hardly surprising that no method exists which would not prefer the closely overlapping Model 2 to the no-where-close Model 1 as the better approximation of Model 3. Nevertheless, decisions based on Model 1 are nearly indistinguishable from those based on the true model [Fig 4B], while Model 2 leads to over-harvesting.  The explanation comes from noticing that the stock size corresponding to the maximum growth rate under Model 1 (the peak of the curve) falls at almost exactly the same stock size as that of the peak growth rate for Model 3.  Meanwhile, the peak of Model 2 occurs at a substantially lower stock size.  While the optimal control solution appears to depend only on a step-ahead forecast accuracy (indeed, the SDP solution method used here takes only step-ahead forecast probabilities as input, [@Marescot2013]), mathematical analysis showed long ago [@Reed1979] that the optimal solution for this problem depends only on keeping the biomass at the value responsible for the maximum growth rate.  This realization is quite general: for most decision problems, simple models will exist under which the optimal decision is the same as it would be for the true model, even when that simple model is wrong in most other ways.

<!-- Intuition for it's not just this example -->
This phenomenon is not unique to sequential decision problems or optimal control solutions.  For example, Management Strategy Evaluation [@Punt2016] in fisheries seeks to evaluate pre-specified strategies rather solve for the best possible strategy through optimal control.  This approach is well justified -- optimal control techniques such as stochastic dynamic programming illustrated here quickly become intractable for more complex models [@Marescot2013] typically used in fisheries stock assessments.  Constraining the search to only strategies that impose a constant mortality (defined as harvest per unit biomass, $F = H/B$) means we only have to evaluate those $N$ strategies each iteration, not solve a sequential decision problem. Doing so, we would find the best constant-mortality solution under Model 1 performs much worse than the best constant-mortality solution under Model 2. Does this mean MSE is not susceptible to this issue?  No, it does not.  It is just as easy to construct an alternative Model 1 with the same properties of leading to nearly optimal decisions while being rejected by any method of iterative learning, such as forecasting or adaptive management.  Just as the optimal control policy for a Gordon-Schaefer (logistic) model depends only on parameter $K$ [@Reed1979], the optimum constant escapement policy depends only on parameter $r$ [@Schaefer1954]. If Model 1 has a value of $r$ such that it happens to match the best possible constant-mortality solution for the (unobserved) true model, then $K$ can be set arbitrarily high, ensuring the any model selection, forecasting, or adaptive management approach would lead away from Model 1 and towards worse-performing models once again. Many optimization problems share this feature in which the optimal policy depends only on a subset or ratio of model parameters, such that it is usually easy to find a similar Model 1. This example also illustrates the importance of decision constraints -- a model that gives very good outcomes when we are free to vary the havest quota each year (the optimal control problem here) may give very poor results under the constraint of constant mortality, and vice versa.

@@

<!-- Not just "get more models" -->
This case cannot be dismissed merely as being dealt a bad hand in having to pick only between Model 1 and Model 2. Model 1 approximates the key feature, giving nearly optimal outcomes.  Previous literature has underscored the importance of Knightian "unknown unknowns" two alternative models examined here fail to fully reflect our uncertainty in the underlying dynamics, and the consequences of underestimating model uncertainty are well understood [@Polasky2011; @Wintle2010]. Had we included the true model in the set of possibilities, the techniques illustrated would have had little difficulty in distinguishing it from Models 1 and 2 after sufficient iterations. In practice, we never have the true model, and even a much more thorough suite of candidate models, such as considering any possible values of $r$, $K$ [Fig 3B], or even among common alternative models with different structural forms, such as Beverton-Holt [@Beverton1957], Ricker [@Ricker1954], or Shepherd [@Shepherd1980], would fail to do any better, since all these models still have symmetric growth rates with a peak growth rate at half the steady-state population size).  Models will always be simpler than reality.  To insist that this issue can and will be avoided by always including more and better models of the process misses the point.

@@

Previous work has acknowledged that iterative learning may not always be feasible and that it may not always be beneficial, but has failed to recognize the possibility that it can potentially be detrimental.  For example, active adaptive management is premised on the observation that future observations may be too uninformative to distinguish between alternatives [@Walters1978].  Others have also noted that even big improvements in predictive accuracy could have negligible improvement on the decision outcomes, which motivates quantifying the Value of Information (VOI) [@Katz1987]. But in both cases, the worse outcome is wasted effort.  In the scenario considered here, the VOI is negative: any learning over the proposed model uncertainty leads to lower expected net utility than not learning.

The only way to avoid this trap is to _forego learning_ with each new observation, to avoid iterative updates.  It may seem that the adaptive management approach fails because posterior probabilities are updated according to Bayes rule. Like iterative forecasting, this favors models which better predict the data, despite the fact that under the adaptive management approach, the overall optimization is conditioned on actual utility and not model fit.  However, this issue is not easily avoided.  For example, so-called greedy optimization techniques that favor actions with higher immediate reward do even worse in this context, since such an algorithm would obviously maximize the immediate harvest and therefore collapse the stock.  Only by comparing the net utility derived from management under model 1 for many iterations to the net utility derived under model 2 after many iterations can we deterime that model 1 leads to better outcomes.  This raises several questions.  Do such methods of learning and reducing model uncertainty ever lead us towards worse long-term outcomes in the real world? If so, How would we know, and what should we do about it?

<!-- **Is it just a unicorn?** -->
Do models which give nearly optimal performance while at the same time making wildly wrong predictions really exist?  It is tempting to argue that Model 1 is a kind of unicorn, a mythical creature existing only in theory. But that is a difficult assertion to prove.  Examples of serious forecasts that widely fail to predict future observations abound wherever forecasts are common, from elections to economics to environmental change [e.g. @Tetlock2015].  If there is no shortage of bad forecasts, then could any of them be useful?  The premise that model need not perfectly capture reality to be useful is at the very heart of modeling.  Model 1 successfully captures the one key feature driving decisions in optimal harvest control problems: the stock size at which the maximum growth rate occurs [Fig 4A].  Capturing only the essential aspects as simply as possible is the goal of any model building exercise.  Thus it should be no surprise that a model can drive good decisions while making poor predictions.  It is also important to note that in many cases, any models being used in decision-making are not necessarily being subjected to the rigorous evaluation and updating steps proposed by adaptive management [@Walters1978] or iterative forecasting [@Dietze2018].  This makes it more likely that such models could persist in practice until now. While iterations that revise this models will no doubt improve decision outcomes in many case, it is worth bearing in mind from the example here that such a connection is not guaranteed.

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  We have seen that the problem cannot be resolved by more data, and is not the result of overfitting. Nor is "creating more models" the answer: when we have a model that is good enough to get optimal results, we cannot always insist on more models. We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations. Approaches such as iterative forecasting or adaptive management that can reduce model uncertainty over time remain promising and important techniques, but because value of information can be negative, we must revise model uncertainty only cautiously.  More importantly, we have seen that  the intuition offered by decision theory [Fig 4] can help us better understand what features of a model are essential to decision outcomes and what are not.  Conversely, models which appear to lead to accurate predictions (like Model 2) can result in outcomes that are far from optimal. Building on such understanding, it may possible to identify strategies for learning that are more agnostic to the details of the model that are not important, or perhaps not reliant on model-based predictions at all. How that is done is an open question.

>>>>

Let us assume $\xi_t(\sigma)$ represents log-normal random noise with mean of unity and log-standard-deviation $\sigma_i$ for each model.
Model 1 is given by $r_1 = 2$, $K_1 = 16$, $\sigma_1 = 0.05$, Model 2 by $r_2 = 0.5$,  $K_2 = 10$, $\sigma = 0.075$ (in dimensionless units).
Having both the larger growth rate and the larger carrying capacity, model 1 is clearly the more optimistic of the two choices.

Selecting between model 1 and model 2 can thus be considered the simplest illustration of the model uncertainty problem.
This is a subset of the more general problem of selecting model parameters, assuming a logistic growth, which itself is a subset of estimating the best structural from (e.g. Ricker, Beverton-Holt, etc).
There is no need to consider these more complicated versions of the model uncertainty problem here, since they all inherit the same issue.
Reducing the model selection problem to these two models simplifies the presentation and will aid intuition at no loss of generality.

@@

The task of deciding whether Model 1 or model 2 would be the better choice for decision making is thus perhaps the simplest example of the much studied issue of model uncertainty that we can pose. 
As in any real world scenario, neither model is the true model, but nevertheless this model set contains a good enough approximation of the true model to make good decisions.
However, any of the well-developed approaches for decision-making under model uncertainty will prefer model 2 over model 1, despite the fact that the optimal policy under model 2 leads to much worse outcomes ecologically and economically.

@@

I illustrate the process of model selection by strictly proper scoring rules using two scenarios.
In Scenario A (passive observation) the fish stock is unharvested and allowed to recover towards carrying capacity (as simulated under our "true" model, Model 3) while comparing the observed stock size in each subsequent time step to the distribution predicted under model 1 and model 2 respectively [Fig 1].
The mean, $\mu_t$ and variance, $\sigma_t$ of the forecast are compared against the true observation $x_t$ using a proper scoring rule given by @Gneiting2007,

@@

In Scenario B (actively harvest), I have first solved for the optimal management strategy for both model 1 and model 2 [Fig 1b].
For small noise and concave functions with linear reward structure this can be done analytically [see proof in @Reed1979], or solved more generally by stochastic dynamic programming [see review by @Marescot2013, details in Appendix].
Under this scenario, replicate simulations of the stock are harvested at each time step using the optimal quota dictated by either model 1 and model 2.
The resulting stock sizes in the time-step following this harvest are once again compared to the probabilities predicted by each model using Eq \eqref{proper}.
Model 2 unequivocally outperforms model 1 in both scenarios of passive observation and active harvest.

Despite the clearly superior predictive accuracy of model 2 in both scenarios, the outcomes from management under model 2 are substantially worse.
We can assess such outcomes in less abstract terms than forecasting skill, such as economic value (in dollars) or the ecological value (unharvested biomass).
In our simple formulation of the decision problem, the "utility" the manager seeks to maximize is simply the economic value (net present value: the discounted sum of all profits from future harvests, Eq \eqref{utility}) of harvested fish.
This formulation ignores any utility provided by fish that are not harvested, beyond their contribution to future potential harvests.
While it is possible to include such contributions directly in the utility function being optimized [e.g. @Halpern2013], even without doing so, model 1 maintains both a higher unharvested biomass and also leads to higher economic returns throughout [Fig 2].

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from model 1 result in a significantly higher fish stock size than under Model 2 (panel A). Economic returns under model 1 are also substantially higher (panel B)"}

@@

In both scenarios, the careful comparison through proper scoring rules has led us to select the worse-performing model.
Crucially, a manager operating under this selection would have little indication that their model was flawed: both future stock sizes and expected harvest yields consistently match model predictions.
Had we been able to include Model 3 in our forecast comparisons, it would equal or outperform the forecasting skill of both model 1 and model 2 (as guaranteed by the theorem of @Gneiting2007), while also matching or out-performing their economic utility (as guaranteed by the theorem of @Reed1979).
In practice, we never have access to the generating model, so it is reasonable to expect model selection to determine the best approximation.
As we see here, the best approximation for forecasting future states does not in fact lead to better outcomes.

@@

Any adaptive management strategy updates posterior distributions over model uncertainty [@Punt2016; @Ludwig1982].
Unfortunately, any such adaptive updating leads to worse outcomes than the equivalent non-adaptive strategy, in which model uncertainty is held fixed.
I illustrate the application of a passive adaptive management strategy to this simple example, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty.
Passive adaptive management for a simple sequential decision problem is straight forward to implement over a discrete set of states and actions using dynamic programming with iterative updates [@Smith1981, example code in Appendix].
To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.
After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct  [Fig 3A].
I compare these results to 
As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating, which achieves a net present value that only 31% that expected under management using model 1 alone [Fig 2].

```{r figure3, fig.height=5, fig.width=8, fig.cap="Adaptive management under model uncertainty.  Solid lines trace the trajectories of the state (fish stock, circles) and action (harvest quota, triangles), under adaptive management (learning).  Dotted lines trace the corresponding trajectories if iterative learning is omitted, leaving the prior belief fixed throughout the simulation (planning).   Color indicates the belief that model 1 is correct (blue), with an intial prior belief of 99%.  Panel A: Management over the two candidate models, Model 1 and Model 2. Within a single iteration of adaptive management, the belief over models switches from a prior belief that heavily favored model 1 to a posterior that favors model 2 with near certainty.  Future iterations reinfornce the belief in model 2, resulting in both depressed harvests and low stock sizes (solid lines).  If no iterative learning updates are performed, stock sizes and realized harvests (and thus economic profit) are both higher. Panel B: given 42 candidate models over a broad range of parameter values, adaptive management quickly reduces the probability on model 1, and substantially underperforms management without learning (dotted lines). While outcomes improve marginally relative to the two-model case (figure A) they remain significantly worse than had no iterative learning been included."}
am <- read_csv("../data/am.csv")
fig3a <- am %>%  filter(time <=20) %>% 
  pivot_longer(c(stock, quota), values_to="stock", names_to="key") %>%
  mutate(key = factor(key, levels = c("stock", "quota"))) %>%
  ggplot(aes(time, stock, shape = key, group = interaction(method,key))) + 
  geom_line(aes(lty=method), col = "gray40", show.legend = FALSE) + 
  geom_point(aes(col = belief), size=2, show.legend = FALSE) +
  scale_colour_gradient(limits = c(0,1), low = pal[4], high = pal[1]) +
  ylab("fish stock") + labs(subtitle="A.")

@@

Increasing the space of possible models to cover a whole plausible range of parameters $r$ and $K$ does little to resolve this problem [Fig 3B]. 
Iterative updates again quickly dismiss the parameter values assumed by model 1, though with more options to choose from, this probability is spead over a range of seemingly plausible candidate models instead of a single alternative model (see Appendix).
While the adaptive management of additional actions and observations slowly narrow this subset of plausible models, decisions based on this uncertainty prevent fish stocks from recovering fully, and realize lower harvests as a result.  
Note that learning under either adaptive management approach (using two models or 42), the decision-maker becomes ever-increasingly convinced that they are using the right model or models.
Future stock sizes fall consistently in the range predicted by the model(s), and consistently outside the range predicted by model 1.
Consequently, each iteration the managers are only more firmly convinced that they are maintaining the fish stock near the biomass that supports maximum sustainable yield, when in fact they are sustaining a harvest regime that is preventing recovery of the stock to the much higher productivity regime which would have been achieved under model 1.

@@

<!-- quick summary of what was demonstrated -->
Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other.
Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.
In both approaches, the manager quickly concludes that model 1 is entirely implausible, while finding that model 2 is remarkably accurate at predicting future values.
This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models spanning a wide range of possible $K$ and $r$ values): which improves only marginally over the outcome of considering only two models [Fig 3B].
Both iterative forecast evaluation and the adaptive management approaches to reducing model uncertainty lead to management becoming trapped in a self-fulfilling prophecy:
holding the fish population at an unproductively low biomass suggested by model 2 while becoming ever more convinced that model is suitable. 
Without access to the true model, none of our methods for model choice or reducing model uncertainty can break this trap.

<!-- Intuition, Figure 4 -->
The reason for model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, model 3.
Plotting the growth rate functions of each model, [Fig 4A], it is hardly surprising that no method exists which would not prefer the closely overlapping model 2 to the no-where-close model 1 as the better approximation of model 3.
Nevertheless, decisions based on model 1 are nearly indistinguishable from those based on the true model [Fig 4B], while model 2 leads to over-harvesting.
The explanation comes from noticing that the stock size corresponding to the maximum growth rate under model 1 (the peak of the curve) falls at almost exactly the same stock size as that of the peak growth rate for model 3.
Meanwhile, the peak of model 2 occurs at a substantially lower stock size.
While the optimal control solution appears to depend only on a step-ahead forecast accuracy (indeed, the SDP solution method used here takes only step-ahead forecast probabilities as input, [@Marescot2013]), mathematical analysis showed long ago [@Reed1979] that the optimal solution for this problem depends only on keeping the biomass at the value responsible for the maximum growth rate.
This realization is quite general: for most decision problems, simple models will exist under which the optimal decision is the same as it would be for the true model, even when that simple model is wrong in most other ways.

<!-- Intuition for it's not just this example -->
This phenomenon is not unique to sequential decision problems or optimal control solutions.
For example, Management Strategy Evaluation [MSE; @Punt2016] in fisheries seeks to evaluate pre-specified strategies rather solve for the best possible strategy through optimal control.
That approach is well justified -- optimal control techniques such as stochastic dynamic programming illustrated here quickly become intractable for more complex models [@Marescot2013] typically used in fisheries stock assessments.
Constraining the search to only strategies that impose a constant mortality (defined as harvest per unit biomass, $F = H/B$) means we only have to evaluate those $N$ strategies each iteration, not solve a sequential decision problem.
Doing so, we would find the best constant-mortality solution under model 1 performs much worse than the best constant-mortality solution under model 2.
Does this mean MSE is not susceptible to this issue?  
No, it does not.
It is just as easy to construct an alternative model 1 with the same properties of leading to nearly optimal decisions while being rejected by any method of iterative learning, such as forecasting or adaptive management.
Just as the optimal control policy for a Gordon-Schaefer (logistic) model depends only on parameter $K$ [@Reed1979], the optimum constant escapement policy depends only on parameter $r$ [@Schaefer1954].
If model 1 has a value of $r$ such that it happens to match the best possible constant-mortality solution for the (unobserved) true model, then $K$ can be set arbitrarily high, ensuring the any model selection, forecasting, or adaptive management approach would lead away from model 1 and towards worse-performing models once again.
Many optimization problems share this feature in which the optimal policy depends only on a subset or ratio of model parameters, such that it is usually easy to find a similar model 1.
This example also illustrates the importance of decision constraints -- a model that gives very good outcomes when we are free to vary the havest quota each year (the optimal control problem here) may give very poor results under the constraint of constant mortality, and vice versa.

@@

<!-- Not just "get more models" -->
This case cannot be dismissed merely as being dealt a bad hand in having to pick only between model 1 and model 2.
Model 1 approximates the key feature, giving nearly optimal outcomes.
Previous literature has underscored the importance of Knightian "unknown unknowns" two alternative models examined here fail to fully reflect our uncertainty in the underlying dynamics, and the consequences of underestimating model uncertainty are well understood [@Polasky2011; @Wintle2010].
Had we included the true model in the set of possibilities, the techniques illustrated would have had little difficulty in distinguishing it from models 1 and 2 after sufficient iterations.
In practice, we never have the true model.
We have also seen that including a much larger suite of possible values of $r$, $K$ [Fig 3B] still leaves us far from the optimal performance of model 1 alone.
Would a more comprehensive grid of $r$ and $K$ values, or the consideration common alternative models with different structural forms, such as Beverton-Holt [@Beverton1957], Ricker [@Ricker1954], or Shepherd [@Shepherd1980], succeed in closing that gap?  
No it would not, as the comparison to the true model makes clear [Fig 4].
Regardless of parameter value, all of these models still have symmetric growth rates with a peak growth rate at half their steady-state population size.
By contrast, the true model has a peak growth rate much closer to the steady-state size.
Because the decision in this problem is driven by the location of the peak growth rate [@Reed1979], any symmetric model which places its peak at the correct stock size also overestimates the carrying capacity, as is the case with our Model 1.
If such additional structurally different models are not enough, then perhaps more complex models would serve?
Models will always be simpler than reality.
To insist that this issue can and will be avoided by always including more and better models of the process misses the point.

@@

The potential for long-term deterioration in management outcomes due to adaptive learning has not been widely acknowledged before.
Certainly the limitations of any modeling approach to improve management outcomes due to the human context of competing interests, imperfect enforcement, etc., is readily appreciated [@Ludwig1993].
The problem here goes deeper, to the limits of models and learning from data in even in a toy example free from other real world challenges.
Previous work has acknowledged that iterative learning may not always be feasible and that it may not always be beneficial [@Polasky2011], but has failed to recognize the possibility that it can potentially be detrimental.
For example, active adaptive management is premised on the observation that future observations may be too uninformative to distinguish between alternatives [@Walters1978].
Others have also noted that even big improvements in predictive accuracy could have negligible improvement on the decision outcomes, which motivates quantifying the Value of Information (VOI) [@Katz1987].
But in both cases, the worse outcome is wasted effort.
In the scenario considered here, the VOI is negative: any learning over the proposed model uncertainty leads to lower expected net utility than not learning.

Worse, absent the counterfactual of non-adaptive management to compare to, this detioriation in outcomes is invisible. 
All the information available to our manager reinforces the conviction that this deteriorated ecological state is in fact the natural, sustainably managed equilibrium of our fish population.
By responding in any way to future observations to reduce model uncertainty, our manager is caught in a trap of self-fulfilling prophecy:
believing ever more firmly in models that make more accurate predictions while driving decisions that prevent the fish stock from recovering.
The only way to avoid this trap is to _forego learning_ with each new observation, to avoid iterative updates.
It may seem that the adaptive management approach fails because posterior probabilities are updated according to Bayes rule.
Like iterative forecasting, this favors models which better predict the data, despite the fact that under the adaptive management approach, the overall optimization is conditioned on actual utility and not model fit.
However, this issue is not easily avoided.
For example, so-called greedy optimization techniques that favor actions with higher immediate reward do even worse in this context, since such an algorithm would obviously maximize the immediate harvest and therefore collapse the stock.
Only by comparing the net utility derived from management under model 1 for many iterations to the net utility derived under model 2 after many iterations can we deterime that model 1 leads to better outcomes.
This raises several questions.
Do such methods of learning and reducing model uncertainty ever lead us towards worse long-term outcomes in the real world? If so, How would we know, and what should we do about it?

<!-- **Is it just a unicorn?** -->
Do models which give nearly optimal performance while at the same time making wildly wrong predictions really exist?
It is tempting to argue that model 1 is a kind of unicorn, a mythical creature existing only in theory. But that is a difficult assertion to prove.
Examples of serious forecasts that widely fail to predict future observations abound wherever forecasts are common, from elections to economics to environmental change [e.g. @Tetlock2015].
If there is no shortage of bad forecasts, then could any of them be useful?  The premise that model need not perfectly capture reality to be useful is at the very heart of modeling. 
Model 1 successfully captures the one key feature driving decisions in optimal harvest control problems: the stock size at which the maximum growth rate occurs [Fig 4A].
Capturing only the essential aspects as simply as possible is the goal of any model building exercise.  Thus it should be no surprise that a model can drive good decisions while making poor predictions.
It is also important to note that in many cases, any models being used in decision-making are not necessarily being subjected to the rigorous evaluation and updating steps proposed by adaptive management [@Walters1978] or iterative forecasting [@Dietze2018].
This makes it more likely that such models could persist in practice until now. While iterations that revise this models will no doubt improve decision outcomes in many case, it is worth bearing in mind from the example here that such a connection is not guaranteed.

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them?
I believe this is an open question, but that our first step must be to recognize it as such.
We have seen that the problem cannot be resolved by more data, and is not the result of overfitting.
Nor is "creating more models" the answer: when we have a model that is good enough to get optimal results, we cannot always insist on more models.
We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations.
Approaches such as iterative forecasting or adaptive management that can reduce model uncertainty over time remain promising and important techniques, but because value of information can be negative, we must be careful not to update models without a keen understanding of what aspects of a model drive a decision [Fig 4].
We have seen that  the intuition offered by decision theory [Fig 4] can help us better understand what features of a model are essential to decision outcomes and what are not.
Conversely, models which appear to lead to accurate predictions (like Model 2) can result in outcomes that are far from optimal.
Building on such understanding, it may possible to identify strategies for learning that are more agnostic to the details of the model that are not important, or perhaps not reliant on model-based predictions at all.
How that is best done in general is an open question.
